beautier: BEAUti for R

### 2017-12-22 article.tex
Rewrite

% Style of code
% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
\usepackage{fancyvrb} 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  language=R,% set programming language
  basicstyle=\ttfamily\small,% basic font style
  commentstyle=\color{gray},% comment style
  numbers=left,% display line numbers on the left side
  numberstyle=\scriptsize,% use small line numbers
  numbersep=10pt,% space between line numbers and code
  tabsize=2,% sizes of tabs
  showstringspaces=false,% do not replace spaces in strings by a certain character
  captionpos=b,% positioning of the caption below
  breaklines=true,% automatic line breaking
  escapeinside={(*}{*)},% escaping to LaTeX
  fancyvrb=true,% verbatim code is typset by listings
  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
  literate={"}{{\texttt{"}}}1{<-}{{$\bm\leftarrow$}}1{<<-}{{$\bm\twoheadleftarrow$}}1
  {~}{{$\bm\sim$}}1{<=}{{$\bm\le$}}1{>=}{{$\bm\ge$}}1{!=}{{$\bm\neq$}}1{^}{{$^{\bm\wedge}$}}1,% item to replace, text, length of chars
  alsoletter={.<-},% becomes a letter
  alsoother={$},% becomes other
  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
  deletekeywords={c}% remove keywords 
}
\lstset{style=mystyle}

@@

\begin{abstract}
  \textbf{1. }
  Here, I present a package, \verb;beautier;, 'BEAUti for R', for the R programming language. \\
  \textbf{2. }
    \verb;beautier; allows for scripted use of the BEAST2 phylogenetics tool, 
    by creating BEAST2 input files from an R function call. \\
  \textbf{3. }
    I describe \verb;beautier; usage, the novel functionality it provides
    compared to BEAUti, and give some minimal examples. \\
  \textbf{4. }
    As \verb;beautier; is free, libre, open-source and designed to be extended, 
    I conclude by describing the current development of the package \\
\end{abstract}

@@

BEAST2 \cite{bouckaert2014beast} is a Bayesian phylogenetics tool.
BEAST2 creates a posterior of jointly-estimated phylonies and model
parameters, from a DNA, RNA or amino acid alignment. BEAST2 can be run
from the command line. To run BEAST2 from the command-line, 
an XML configuration file is needed, that contains the alignment and model parameters.

@@

\verb;beautier; is written in the R programming language \cite{R}.
\verb;beautier; creates the BEAST2 input files from an R function call.

\verb;beautier;'s main function is \verb;create_beast2_input_file;, which creates
an BEAST2 input file.

\verb;create_beast2_input_file; needs at least the name of a FASTA file containing a DNA alignment
and a name for the to-be-created output file. This interface follows BEAUti's default settings.
Per alignment, a site model, clock model and tree priors can be chosen.
Multiple alignments can be used, each with its own (unlinked) site model, clock model and tree prior.

@@

In total, \verb;beautier; has 59 exported functions to create
a BEAST2 configuration file. \verb;beautier; is an alternative for a majority of BEAUti use cases.
\verb;beautier;s future extensions (also: road map) can be found on its GitHub.

\verb;beautier;, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

\verb;beautier; has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

\verb;beautier; does not support the full functionality of BEAUti. Considering
the size, age and number of plugins, this would be close to impossible.
To compensate for this, an extensible software architecture is used.

\verb;beautier; has minimal support for calling BEAST2 from within R and does
so for testing purposes only.

\verb;beautier; has minimal support for parsing and interpreting BEAST2 output files,
for that the \verb;beastier; \cite{beastier} package is recommended.

@@

\begin{lstlisting}[language=R, caption=Example with fixed crown age, label=lst:fixed_crown_age, floatplacement=H]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = fasta_to_phylo(
    fasta_filename = "alignment.fas", 
    crown_age = 15
  )
)
\end{lstlisting}

@@

\verb;beautier; is dependent on multiple packages, which are 
\verb;APE; \cite{APE}, 
\verb;devtools; \cite{devtools},
\verb;geiger; \cite{GEIGER},
\verb;ggplot2; \cite{ggplot2},
\verb;knitr; \cite{knitr},
\verb;phangorn; \cite{phangorn},
\verb;RBeast; \cite{RBeast},
\verb;rmarkdown; \cite{rmarkdown},
\verb;seqinr; \cite{seqinr},
\verb;stringr; \cite{stringr},
\verb;testit; \cite{testit} and 
\verb;TreeSim; \cite{TreeSim}.

\verb;beautier;'s documentation is extensive, yet concise. All funnctions are documented
in the package's internal documentation. For quick use, each exported function shows a minimal example. 
For easy exploration, each exported function's documentation links to related functions.
Additionally, \verb;beautier; has a vignette that demonstrates in a longer form how
to use it. The integrity of this documentation is tested each time the package is built by Travis CI.
The documentation on the GitHub helps to get started, with a dozen examples 
of a BEAUti screenshot and the equivalent \verb;beautier; code.

>>>>

% Style of code
% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
\usepackage{fancyvrb} 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  language=R,% set programming language
  basicstyle=\ttfamily\small,% basic font style
  commentstyle=\color{gray},% comment style
  numbers=left,% display line numbers on the left side
  numberstyle=\scriptsize,% use small line numbers
  numbersep=10pt,% space between line numbers and code
  tabsize=2,% sizes of tabs
  showstringspaces=false,% do not replace spaces in strings by a certain character
  captionpos=b,% positioning of the caption below
  breaklines=true,% automatic line breaking
  escapeinside={(*}{*)},% escaping to LaTeX
  fancyvrb=true,% verbatim code is typset by listings
  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
  literate={"}{{\texttt{"}}}1{<-}{{$\bm\leftarrow$}}1{<<-}{{$\bm\twoheadleftarrow$}}1
  {~}{{$\bm\sim$}}1{<=}{{$\bm\le$}}1{>=}{{$\bm\ge$}}1{!=}{{$\bm\neq$}}1{^}{{$^{\bm\wedge}$}}1,% item to replace, text, length of chars
  alsoletter={.<-},% becomes a letter
  alsoother={$},% becomes other
  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
  deletekeywords={c}% remove keywords 
}
\lstset{style=mystyle}

@@

\begin{abstract}
  \textbf{1. }
  Here, we present a package, \verb;beautier;, 'BEAUti for R', for the R programming language. \\
  \textbf{2. }
    \verb;beautier; allows for scripted use of the BEAST2 phylogenetics tool, 
    by creating BEAST2 input files from an R function call. \\
  \textbf{3. }
    I describe \verb;beautier; usage, the novel functionality it provides
    compared to BEAUti, and give some minimal examples. \\
  \textbf{4. }
    As \verb;beautier; is free, libre, open-source and designed to be extended, 
    I conclude by describing the current development of the package \\
\end{abstract}

@@

BEAST2 \cite{bouckaert2014beast} is a Bayesian phylogenetics tool.
BEAST2 creates a posterior of jointly estimated phylonies and model
parameters, from a DNA, RNA or amino acid alignment. BEAST2 can be run
from the command line. To run BEAST2 from the command-line, 
an XML configuration file is needed, that contains the alignment and model parameters.

@@

\verb;beautier; is written in the R programming language \cite{R}.
\verb;beautier; creates the BEAST2 input files from an R function call,
in a similar way that BEAUti does.

\verb;beautier;'s main function is \verb;create_beast2_input_file;, which creates
an BEAST2 input file. \verb;create_beast2_input_file; needs at least the name of a FASTA file containing a DNA alignment
and a name for the to-be-created output file. This interface follows BEAUti's default settings.
Per alignment, a site model, clock model and tree priors can be chosen.
Multiple alignments can be used, each with its own (unlinked) site model, clock model and tree prior.

@@

In total, \verb;beautier; has 59 exported functions to create
a BEAST2 configuration file. \verb;beautier; is an alternative for a majority of BEAUti use cases.
\verb;beautier; does not support the full functionality of BEAUti. Considering
the size, age and number of plugins, this would be close to impossible.
To compensate for this, an extensible software architecture is used.
\verb;beautier;s future extensions can be found on its GitHub.

BEAUti assumes that a phylogeny has a crown age that needs to be jointly estimated
with the phylogeny and other parameters. BEAUti does not allow for fixing
a phylogeny's crown age. Before \verb;beautier;, one needs to manually edit the BEAST2 
XML configuration file, which is prone to errors. \verb;beautier;, allow easy
fixing of phylogenies' crown ages.

\verb;beautier; has only internal support for calling BEAST2 from within R and does
so for testing purposes only. \verb;beautier; has minimal support for parsing and interpreting BEAST2 output files,
for that the \verb;beastier; \cite{beastier} package is recommended.

@@

\begin{lstlisting}[language=R, caption=Example with fixed crown age, label=lst:fixed_crown_age, floatplacement=H]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
  fixed_crown_ages = TRUE,
  initial_phylogenies = fasta_to_phylo(
    fasta_filename = "alignment.fas", 
    crown_age = 15
  )
)
\end{lstlisting}

@@

\verb;beautier; is dependent on multiple packages, which are 
\verb;APE; \cite{APE}, 
\verb;devtools; \cite{devtools},
\verb;geiger; \cite{GEIGER},
\verb;ggplot2; \cite{ggplot2},
\verb;knitr; \cite{knitr},
\verb;phangorn; \cite{phangorn},
\verb;beastier; \cite{beastier},
\verb;rmarkdown; \cite{rmarkdown},
\verb;seqinr; \cite{seqinr},
\verb;stringr; \cite{stringr},
\verb;testit; \cite{testit} and 
\verb;TreeSim; \cite{TreeSim}.

\verb;beautier;'s documentation is extensive, yet concise. All functions are documented
in the package's internal documentation. For quick use, each exported function shows a minimal example. 
For easy exploration, each exported function's documentation links to related functions.
Additionally, \verb;beautier; has a vignette that demonstrates in a longer form how
to use it. The integrity of this documentation is tested each time the package is built by Travis CI.
The documentation on the GitHub helps to get started, with a dozen examples 
of a BEAUti screenshot and the equivalent \verb;beautier; code.
